### 2018-12-16 paper/paper.tex
Formatting fixes. Make line graphs pareable in B&W
Small grammar corrections

    \begin{multicols*}{2}
        \section{Introduction}
            The classification of spam emails is an important task in today's society,
            as we receive potentially hundreds of them a year which clutter
            an already full inbox. These emails also pose a threat to the uninformed
            as many ask for personal information, such as street addresses and banking details.
            Traditionally emails have been filtered through rule-based systems that
            could be reverse engineered to still allow for spam email to get through.
            Some newer filters take advantage of machine learning techniques and use
            them in conjunction with the traditional method which leads to good results,
            but still some emails get through. Some newer classifiers may hold the
            key to filtering out even the most engineered emails.

@@

            Next we used BeautifulSoup's HTML parser to interpret any HTML data that
            was in an email \cite{beautiful_soup}. We queried the document structure for the \textit{href}
            attribute because it indicates a link. Not only an \textit{a} tag can include
            a link, tags such as an \textit{img} tag can also contain one. Potential attackers
            may also use an \textit{a} tag to make a link look like it is going somewhere else,
            such as having the text \textit{http://google.com} a clickable link to
            \textit{http://mygooglephishingsite.com}, but we didn't see this practice 
            occurring in this dataset so we ignored this potential feature. We also looked at the
            amount of linked pictures, as an attacker may try to get credibility with it's prey
            by introducing fancy graphics. Lastly we extracted the raw content from the HTML,
            ignoring any structure that was present for tokenization.

            The tokenization step was done with NLTK's word and sentence tokenizer.
            The sentence tokens were used for some metrics such as average words per
            sentence and could be used in the future for the analysis of sentence
            structure \cite{nltk}. The word tokens were stripped of punctuation such as '.',
            '?', '(' and ')', which would litter our unigrams. Next we removed stopwords
            from the provided list of English stopwords in NLTK's corpus. Removing
            these words are necessary as they provide no direct meaning when it comes
            to frequency analysis of the documents and would just provide more noise
            in the classifier. Lastly we
            stemmed the words using NLTK's Porter Stemmer with their extensions to
            Martin Porter's algorithm \cite{porter1980algorithm}. This allowed use to remove suffixes such as '-s',
            which would group `business' and `businesses' separately.

@@

            \begin{Figure}
                \centering
                \includegraphics[scale=0.5]{figures/extract.png}
                \captionof{figure}{Extraction Process}\label{fig:extrac}
            \end{Figure}

            The first step of extraction was
            to find words that meant the most to the corpus which we defined as
            the body of the email. To accomplish this
            we used the tf-idf algorithm implemented in scikit-learn. This algorithm helps
            weigh down words that are too common, and bring up words that are important
            to the corpus. Next we split the documents into spam and ham before analysis.
            This was done because this algorithm is still based on frequency and the
            ratio of ham to spam was not 1:1. Characteristic words to spam emails such
            as ``free'' wouldn't show up because they're not prominent in ham emails.

            \begin{minipage}{0.48\columnwidth}
                \includegraphics[width=\textwidth]{figures/spam_wc}
                \captionof{figure}{Spam}\label{fig:spam_wc}
            \end{minipage}
            \begin{minipage}{0.48\columnwidth}
                \includegraphics[width=\textwidth]{figures/ham_wc}
                \captionof{figure}{Ham}\label{fig:ham_wc}
             \end{minipage}

             Figures \ref{fig:spam_wc} and \ref{fig:ham_wc} show
             the most prominent 50 words in the corpus. Most the the words in the
             ham word cloud relate to business operations asking someone to do about
             task or getting times and dates of events. When you look at the spam
             word cloud you see a few similar words such as ``mail'' or ``get''.
             But in general you see words that you would associate with someone trying
             to get you to send them money or buy a product. Words like ``money'',
             ``offer'' and ``order'' are the epitome of this.

@@

            We also examined the effect of tuning the number of feature words chosen by the TF-IDF 
            extractor. The number of feature words chosen for initial analysis was 50, but subsequent 
            tests ranged from 50 to 1050 in steps of 200.
        \section{Results}
            Classifiers were evaluated on precision and recall on the spam class. Overall, precision is
            weighted more heavily than recall in the evaluation because it would likely be more disruptive 
            to classify a ham email as spam than to let in an occasional spam message. Subsequently, the 
            classifiers were also subjected to various feature word counts to observe the effect.
            \subsection{General Precision and Recall}
                Overall, the classifiers performed well with just features extracted from the email body.
                Precision and recall comparison can be found in \autoref{fig:precision_recall} with direct 
                number comparison in \autoref{tab:precision_recall}.

@@

                \begin{Figure}
                    \centering
                    \includegraphics[width=\linewidth]{figures/precision_recall.png}
                    \captionof{figure}{Precison and Recall Comparison}
                    \label{fig:precision_recall}
                \end{Figure}

@@

            \subsection{Effect of Number of Feature Words} 
                When increasing the number of feature words, precision increased dramatically for some classifiers. 
                Specifically, Naive Bayes and the SVM gained a precision increase of 10\% and 8\% respectively (See \autoref{fig:precision_word_effect}).
                However, the number of feature words did not seem to effect recall as much as precision (See \autoref{fig:recall_word_effect}).
                Only one classifier gained over 3\% precision from 50 feature words. Naive Bayes gained 10\% recall when increasing 
                the number of feature words from 50 to 250. This gain dropped off as the number of feature words increased.

                \begin{Figure}
                    \centering
                    \includegraphics[width=\linewidth]{figures/recall_wv_effect.png}
                    \captionof{figure}{Effect of the number of feature words on classifier recall}
                    \label{fig:recall_word_effect}
                \end{Figure}

                \begin{Figure}
                    \centering
                    \includegraphics[width=\linewidth]{figures/precision_wv_effect.png}
                    \captionof{figure}{Effect of the number of feature words on classifier precision}
                    \label{fig:precision_word_effect}
                \end{Figure}

@@

            This paper has shown that the usage of machine learning alone on email body text 
            along with a few other features can perform the spam classification task well. 
            However, our metrics are lower
            than the other papers we are comparing against ourselves. Adding more features
            that aren't word based may help even more with performance and also increasing the
            size of the dataset would help as well. We also recognize the limitations of
            this implementation lending itself to only classify emails from the early 2000s, as
            we did no testing on modern email data which may have a distribution that doesn't
            work well with our implementation.
        \nocite{*}
        \bibliography{paper} 
        \bibliographystyle{alpha}
    \end{multicols*}
\end{document}

>>>>

    \begin{multicols}{2}
        \section{Introduction}
            The classification of spam emails is an important task in today's society,
            as we receive potentially hundreds of them a year which clutter
            an already full inbox. These emails also pose a threat to the uninformed
            as many ask for personal information, such as street addresses and banking details.
            Traditionally emails have been filtered through rule-based systems that
            could be reverse engineered to still allow for spam email to get through.
            Some newer filters take advantage of machine learning techniques and use
            them in conjunction with the traditional method which leads to good results,
            but still some emails get through. Some newer classifiers may hold the
            key to filtering out even the most engineered emails.

@@

            Next we used BeautifulSoup's HTML parser to interpret any HTML data that
            was in an email \cite{beautiful_soup}. We queried the document structure for the \textit{href}
            attribute because it indicates a link. Not only an \textit{a} tag can include
            a link, tags such as an \textit{img} tag can also contain one. Potential attackers
            may also use an \textit{a} tag to make a link look like it is going somewhere else,
            such as having the text \textit{http://google.com} a clickable link to
            \textit{http://mygooglephishingsite.com}, but we didn't see this practice 
            occurring in this dataset so we ignored this potential feature. We also looked at the
            amount of linked pictures, as an attacker may try to get credibility with their prey
            by introducing fancy graphics. Lastly we extracted the raw content from the HTML,
            ignoring any structure that was present to prepare for tokenization.

            The tokenization step was done with NLTK's word and sentence tokenizer.
            The sentence tokens were used for some metrics such as average words per
            sentence and could be used in the future for the analysis of sentence
            structure \cite{nltk}. The word tokens were stripped of punctuation such as '.',
            '?', '(' and ')', which would litter our unigrams. Next we removed stopwords
            from the provided list of English stopwords in NLTK's corpus. Removing
            these words are necessary as they provide no direct meaning when it comes
            to frequency analysis of the documents and would just provide more noise
            in the classifier. Lastly, we
            stemmed the words using NLTK's Porter Stemmer with their extensions to
            Martin Porter's algorithm \cite{porter1980algorithm}. This allowed use to remove suffixes such as '-s',
            which would group `business' and `businesses' separately.

@@

            \begin{Figure}
                \centering
                \includegraphics[width=\linewidth]{figures/extract.png}
                \captionof{figure}{Extraction Process}\label{fig:extrac}
            \end{Figure}

            The first step of extraction was
            to find words that meant the most to the corpus which we defined as
            the body of the email. To accomplish this
            we used the tf-idf algorithm implemented in scikit-learn. This algorithm helps
            weigh down words that are too common, and bring up words that are important
            to the corpus. Next we split the documents into spam and ham before analysis.
            This was done because this algorithm is still based on frequency and the
            ratio of ham to spam was not 1:1. Characteristic words to spam emails such
            as ``free'' wouldn't show up because they're not prominent in ham emails.\\\\
            \noindent\begin{minipage}{0.48\columnwidth}
                \includegraphics[width=\textwidth]{figures/spam_wc}
                \captionof{figure}{Spam}\label{fig:spam_wc}
            \end{minipage}
            \begin{minipage}{0.48\columnwidth}
                \includegraphics[width=\textwidth]{figures/ham_wc}
                \captionof{figure}{Ham}\label{fig:ham_wc}
             \end{minipage}

             Figures \ref{fig:spam_wc} and \ref{fig:ham_wc} show
             the most prominent 50 words in the corpus. Most of the words in the
             ham word cloud relate to business operations asking someone to do about
             task or getting times and dates of events. When you look at the spam
             word cloud you see a few similar words such as ``mail'' or ``get''.
             But in general you see words that you would associate with someone trying
             to get you to send them money or buy a product. Words like ``money'',
             ``offer'' and ``order'' are the epitome of this.

@@

            We also examined the effect of tuning the number of feature words chosen by the TF-IDF 
            extractor. The number of feature words chosen for initial analysis was 50, but subsequent 
            tests ranged from 50 to 1050 in steps of 200.

            \section{Results}
            Classifiers were evaluated on precision and recall on the spam class. Overall, precision is
            weighted more heavily than recall in the evaluation because it would likely be more disruptive 
            to classify a ham email as spam than to let in an occasional spam message. Subsequently, the 
            classifiers were also subjected to various feature word counts to observe the effect.
    \end{multicols}
        \begin{Figure}
            \centering
            \includegraphics[width=\linewidth]{figures/precision_recall.png}
            \captionof{figure}{Precison and Recall Comparison}
            \label{fig:precision_recall}
        \end{Figure}
    \begin{multicols}{2}
            \subsection{General Precision and Recall}
                Overall, the classifiers performed well with just features extracted from the email body.
                Precision and recall comparison can be found in \autoref{fig:precision_recall} with direct 
                number comparison in \autoref{tab:precision_recall}.

@@

@@

    \end{multicols}
        \begin{Figure}
            \centering
            \includegraphics[width=\linewidth]{figures/recall_wv_effect.png}
            \captionof{figure}{Effect of the number of feature words on classifier recall}
            \label{fig:recall_word_effect}
        \end{Figure}

        \begin{Figure}
            \centering
            \includegraphics[width=\linewidth]{figures/precision_wv_effect.png}
            \captionof{figure}{Effect of the number of feature words on classifier precision}
            \label{fig:precision_word_effect}
        \end{Figure}
    \begin{multicols}{2}
            \subsection{Effect of Number of Feature Words} 
                When increasing the number of feature words, precision increased dramatically for some classifiers. 
                Specifically, Naive Bayes and the SVM gained a precision increase of 10\% and 8\% respectively (See \autoref{fig:precision_word_effect}).
                However, the number of feature words did not seem to effect recall as much as precision (See \autoref{fig:recall_word_effect}).
                Only one classifier gained over 3\% precision from 50 feature words. Naive Bayes gained 10\% recall when increasing 
                the number of feature words from 50 to 250. This gain dropped off as the number of feature words increased.

@@

            This paper has shown that the usage of machine learning alone on email body text 
            along with a few other features can perform the spam classification task well. 
            However, our metrics are lower
            than the other papers we are comparing against ourselves. Adding more features
            that aren't word based may help even more with performance and also increasing the
            size of the dataset would help as well. We also recognize the limitations of
            this implementation lending itself to only classify emails from the early 2000s, as
            we did no testing on modern email data which may have a distribution that doesn't
            work well with our implementation.
        \nocite{*}
        \bibliography{paper} 
        \bibliographystyle{alpha}
    \end{multicols}
\end{document}
